Are you moving predictably?

### 2018-07-15 dokumentation/MiriamClustering.tex
koen hat korrigiert

In \Cref{fig:OrgDist} is the result to see of the first try. \Cref{fig:OrgSt} shows the strata distribution as pie chart and \Cref{fig:OrgCl} the resulted cluster distribution. It can already been seen, that the distributions are not similar. For more steps the outcome is analogous.

The next idea is: are 6 cluster too fine? Based on the grouped strata we look at 3 clusters.

@@

\Cref{fig:OrgDist3Cl} looks promising, so we had a look at the cluster distribution in the 3 grouped stratas, visualized \Cref{fig:Groupdist}. It can be seen, that all clusters appear in all grouped strata.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\textwidth]{vectorClustering31000.PNG}
\caption{1000 steps clustering in 3 clusters on maximal stratified person dataset}
\label{fig:1000vect3}
\end{figure}

\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets we generated creates similar results.

%PCA?

>>>>

\Cref{fig:OrgDist} is the result of the first try. \Cref{fig:OrgSt} shows the strata distribution as pie chart and \Cref{fig:OrgCl} the resulted cluster distribution. It can already been seen, that the distributions are not similar. For more steps the outcome is analogous.

The next idea is: are 6 cluster too specific? Based on the grouped strata we look at 3 clusters.

@@

\Cref{fig:OrgDist3Cl} looks promising, so we had a look at the cluster distribution in the 3 grouped stratas, visualized in \Cref{fig:Groupdist}. It can be seen, that all clusters appear in all grouped strata.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.69\textwidth]{vectorClustering31000.PNG}
\caption{1000 steps clustering in 3 clusters on maximal stratified person dataset}
\label{fig:1000vect3}
\end{figure}

\Cref{fig:1000vect3} shows clearly, that again no correlation can be found. Applying the procedure to the different datasets we generate similar results.

%PCA?
To save time we apply the PCA process on the \textit{original data} without strata and a variance
threshold of 0.95. We get two Attributes, so we can assume, that the data is strongly correlated. As expected the computing time decreases a lot for the whole process of clustering.

\begin{figure}[H]
\vspace*{-1em}
\centering
\begin{subfigure}{.3\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidStrata.PNG}
  \caption{Strata}
  \label{fig:PCAOrgSt}
\end{subfigure}%
\begin{subfigure}{.3\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster.PNG}
  \caption{Cluster}
  \label{fig:PCAOrgCl}
\end{subfigure}
\begin{subfigure}{.3\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{ClusterPCAOrigRapidCluster2Cluster.PNG}
  \caption{3Cluster}
  \label{fig:PCAOrgCl3}
\end{subfigure}
\caption{Distribution of original data}
\label{fig:PCAOrgDist}
\vspace*{-2em}
\end{figure}

For a clustering of 6 clusters we get a distribution, which can be seen in \Cref{fig:PCAOrgDist}, similar to the one we get from the k-means process with the same configurations. 
For a clustering of 3 clusters we again get results, that are unclear(\Cref{fig:PCAOrgCl3}).

Using different dataset results in comparable unrewarding results.
\vspace*{-2em}

### 2018-07-15 dokumentation/conclusion.tex
koen hat korrigiert

	\section{Discussion} \label{sec: discussion}
%	\begin{itemize}
%		\item repr. working day
%		\item over/under std's
%		\item indiv. of lifestyle
%		\item inv. behaviour in same strata
%	\end{itemize}
	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we though about various aspects having an impact on the movement. We state 4 main influencing aspects, explaining the variance throughout the data.

	\subsection{Representativity of the day}
	The day the data was collected on (c.f. \Cref{sec: introduction}) is not mentioned. Therefore we cannot infer, that it is representative. The people, asked to enter their movement, could have a exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day. \\
	Also the day itself is not mentioned, so we do not have information, if it was even a working day or a weekend. This influences the behavior drastically.

@@

	\subsection{Individuality of lifestyle}
	Obviously, people are individual in their way of life. So there are people, with enough money to buy for example a car, but refuse to in order to reduce CO$_2$ emission, or simply do not like driving. On the other end of the spectrum, some people, who have little money, still own and drive a car daily in order to go to work, since they cannot afford an other apartment.\\
	One can think of multitude scenarios of people behaving different caused by their individuality.

	\subsection{Inverse behavior in a strata}
	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
	\begin{enumerate}
		\setlength{\itemindent}{1cm}
		\item[1. Group]
		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
		\item[2. Group]
		People, who are rich just by birth, which do not have the necessity to work or move at all. They could possibly stay home and do not have to leave at all.
	\end{enumerate} 
	Those two groups are completely opposite, but still belong to the same strata. Now also considering strata one, we can find the exact same movement behavior in there as well. Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work. Also there are people, who do not move at all, because of the same reason.

	So we can see that based on the information we have, it is very unlikely to have a clear correlation between the given data and the strata. However, given more information about the previously mentioned aspects and more entries in general, there could be some kind of correlation to be found.\\
	Future work could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering. Also one could use different network visualization tools in order to infer patterns. An example would be to take the origin (destination) sectors as nodes, an edge, if there is a movement between, directed or undirected, and a different thickness or color gradient, based on the number of this edge being taken.\\

>>>>

	\section{Discussion} \label{sec: discussion}
%	\begin{itemize}
%		\item repr. working day
%		\item over/under std's
%		\item indiv. of lifestyle
%		\item inv. behaviour in same strata
%	\end{itemize}
	Regarding the results, of not predicting the strata and also not observing any meaningful clusters throughout the data in given and stratified form, we thought about various aspects having an impact on the movement. We state 4 main influencing aspects, explaining the variance throughout the data.

	\subsection{Representativity of the day}
	The day the data was collected on (c.f. \Cref{sec: introduction}) is not mentioned. Therefore we cannot know, if it is representative. The people, asked to enter their movement, could had an exceptional day, like a vacation day, a doctors appointment or a broken car and therefore behave different from a usual day. \\
	Also the day itself is not mentioned, so we cannot determine, if it was even a working day or a weekendday. This influences the behaviour drastically.

@@

	\subsection{Individuality of lifestyle}
	Obviously, people are individual in their way of life. So there are people, with enough money to buy for example a car, but refuse to do so in order to reduce CO$_2$ emission, or simplydislike driving. On the other end of the spectrum, some people, who have little money, still own and drive a car daily in order to go to work, since they cannot afford an other apartment.\\
	One can think of multitude scenarios of people behaving different caused by their individuality.

	\subsection{Inverse behavior in a strata}
	As an example, take a look at the strata 6, which denotes the richest people considered. Their we have inter alia two groups: 
	\begin{enumerate}
		\setlength{\itemindent}{1cm}
		\item[1. Group]
		Hard working people, laboring 60+ hours per weak to earn their money. One can imagine that they have to move quite a lot since they are always busy.
		\item[2. Group]
		People, who are rich by birth, which do not have the necessity to work or move at all. They could possibly stay home and do not have to leave at all.
	\end{enumerate} 
	Those two groups are completely opposite, but still belong to the same strata. Now also considering strata 1, we can find the exact same movement behaviour in there as well. Within the poorest there are people that are wandering around the town via feet or bike, since they have no objective, like work. Also there are people, who do not move at all, because of the same reason.

	So we can see that based on the information we have, it is very unlikely to have a clear correlation between the given data and the strata. However, given more information about the previously mentioned aspects and more entries in general, a kind of correlation could be found.\\
	Future research could include a component analysis of the decision tree and neural network in order to ascertain the influencing parts and therefore improve the data gathering. Also one could use different network visualization tools in order to infer patterns. An example would be to take the origin (destination) sectors as nodes, an edge, if there is a movement between, directed or undirected, and a different thickness or color gradient, based on the number of this edge being taken.\\

### 2018-07-15 dokumentation/introduction.tex
koen hat korrigiert

\section{Introduction} \label{sec: introduction}
	% could add more in depth 
	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the collected data? In particular is it possible to determine the wealth of a person, only given movements of a single day?
	For this we considered the dataset stated in \cite{rich_do_not_rise_early}. There we have a set of 124979 rows of movement data of various persons from the Columbian town Medellín, which is the second largest Colombian town with an estimated population of 2.5 million as of 2017 \cite{population_number} . All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
	The data entries consist of data about the movement, like endpoints, length, and also some meta parameters like gender, age or the so-called strata of the person. The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs. Those costs are defined by Colombians laws, which classifies households to regulate the access to public utility services, having as a result six socio-economic strata \cite{rich_do_not_rise_early}.\\
	Our goal is to ascertain, if there is a correlation between the movements and the strata, in order to be able to predict the strata based on the movements. This is done in two steps:\\
	First, by clustering using the k-means algorithm, considering different distance measures, with optionally including the principal component analysis (PCA).\\
	And second, a decision tree and neural net were trained and tested.

>>>>

\section{Introduction} \label{sec: introduction}
	% could add more in depth 
	Within the time of Industry 4.0 and various data sources the question arises, if one can define who we are by the collected data? In particular is it possible to determine the wealth of a person, only given movements of one single day?
	For this we considered the dataset stated in \cite{rich_do_not_rise_early}. There we have a set of 124979 rows of movement data of various persons from the Columbian town Medellín, which is the second largest Colombian town with an estimated population of 2.5 million as of 2017 \cite{population_number} . All the data was collected at a single day, with possibly multiple entries referring to the same person.\\
	The data entries consist of data about the movement, like endpoints, length, and also some meta parameters like gender, age or the so-called strata of the person. The strata defines the socio-economic group, reflecting the affluence and therefore impose the ancillary costs. Those costs are defined by Colombians laws, which classifies households to regulate the access to public utility services, having as a result six socio-economic strata \cite{rich_do_not_rise_early}.\\
	Our goal is to ascertain, if there is a correlation between the movements and the strata, in order to be able to predict the strata based on the movements. This is done in two steps:\\
	First, by clustering using the k-means algorithm, considering different distance measures, with optionally including the principal component analysis (PCA).\\
	And second, a decision tree and neural net by training and testing.

### 2018-07-15 dokumentation/observation.tex
koen hat korrigiert

\section{Conclusion} \label{sec: observation}
	%	TODO (1 page)
	%	TODO: restructure. Have like 3 parts
	%	The greater the population, the more the borders between the stratas blur. 
	As an overall result we get, that we can not determine the strata based on the information we have.
	Using various datasets (c.f. \Cref{sec: proprocessing}), we witnessed, that having equally distributed datasets leads to an overall higher accuracy. This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f. \Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
	During clustering we observed, that using a different distance measure formula would not lead to a huge difference. Also, we detected that reducing the numbers of clusters leads to better results, but decreases the significance of the statement, that could be made.\\
	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted as mentioned above.\\
	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions. Therefore we see, the more data we have, the higher is the variance within the stratas themselves. We cannot draw some kind of lines to distinguish between different stratas. Data points, which were outliers in the smaller sets, are now not considered to be outliers, because many others have the same characteristics. So the lines, we were able to draw for smaller datasets, are blurring.\\ %TODO: letzter satz doof	-> hihi

>>>>

\section{Conclusion} \label{sec: observation}
	%	TODO (1 page)
	%	TODO: restructure. Have like 3 parts
	%	The greater the population, the more the borders between the stratas blur. 
	As an overall result we get, that we can not determine the strata based on the information we have.
	Using various datasets (c.f. \Cref{sec: proprocessing}), we witnessed, that having equally distributed datasets leads to an overall higher accuracy. This rules out the bias observed in the original data, where strata 2 and 3 are very dominant (c.f. \Cref{table: distribution normal}), but also reduces the set size from originally 124978 to $6*2038=12228$ entries equally distributed over all stratas.\\
	During clustering we observed, that using a different distance measure formula would not lead to a huge difference. Also, we detected that reducing the numbers of clusters leads to better results, but decreases the significance of the statement, that could be made.\\
	Using a different neural net architecture will most likely not chance the accuracy value drastically. As stated, more complex nets need more training data, which is restricted for reasons mentioned above.\\
	What we can observe is, that using the stratified vectors, we are able to increase the performance and accuracy, but still have no sufficient predictions. Therefore we see, the more data we have, the higher the variance within the stratas. We cannot draw lines to distinguish between different stratas. Data points, which were outliers in the smaller sets, are now not considered to be outliers, because many others have the same characteristics. So the lines, we were able to draw for smaller datasets, are blurring.\\ %TODO: letzter satz doof	-> hihi

### 2018-07-15 dokumentation/preprocessing.tex
Schlaf ist nur für schwache Menschen...
koen hat korrigiert

\subsection{Preprocessing}\label{sec: proprocessing}
	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
	We observe that even though we have 124979 individual lines defining a movement, there is one line defining a \texttt{NotANumber}-exception and therefore gets neglected for further usage.	\\
	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets used by our clustering and neural net approaches.\\
	We observe the following distribution over the whole dataset:\\

	\vspace*{-3em}
	\begin{figure}[H]
		\centering		
		\setlength\tabcolsep{.2cm}
		\begin{tabular}{c|ccccccc}
			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
			abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
		\end{tabular}
		\vspace*{-2.5em}
		\label{table: distribution normal}
	\end{figure}
	There is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
	Furthermore we have to make sure that two different data points, which belong to the very same person, are assigned to the same cluster. To do so, we compute the value \texttt{ID}, which identifies each person and can be used to combine movements that are considered to be from the same person. I.e. two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender. This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.\\

>>>>

\subsection{Preprocessing}\label{sec: proprocessing}
	In order to classify the given data into smaller test sets or mask different aspects, we have to perform some analysis.\\
	We observe that even though we have 124979 individual lines defining a movement, there is 1 line defining a \texttt{NotANumber}-exception and therefore gets excluded from further usage.	\\
	We provide the \texttt{testDataGenerator} python script. Through flags and input arguments, the script is able to create all test sets used by our clustering and neural net approaches.\\
	We observe the following distribution over the whole dataset:\\

	\vspace*{-3em}
	\begin{figure}[H]
		\centering		
		\setlength\tabcolsep{.2cm}
		\begin{tabular}{c|ccccccc}
			strata &  1   &   2   &   3   &  4   &  5   &  6   & $\Sigma$ \\ \hline
			abs   & 6963 & 52265 & 49404 & 8772 & 5536 & 2038 &  124978  \\
			\%   & 5.57 & 41.82 & 39.53 & 7.02 & 4.43 & 1.63 &   100
		\end{tabular}
		\vspace*{-2.5em}
		\label{table: distribution normal}
	\end{figure}
	There is an upper bound on equal distribution through strata 6. It has at most 2038 individual elements.
	Furthermore, we have to make sure that two different data points, which belong to the very same person, are assigned to the same cluster. To do so, we compute the value \texttt{ID}, which identifies each person and can be used to combine movements that are considered to be from the same person. I.e. two movements correspond with the very same person, if and only if they are consecutive in the original dataset and have the same strata, age and gender. This approach is taken since the surveys are concatenated sequentially and it is unlikely, that multiple consecutive movements with same strata, age, gender belong to two different persons.\\

### 2018-07-15 dokumentation/submission.tex
koen hat korrigiert

	%
	%	\title{Clustering Analysis of Mobility Data}
	\title{Are you moving predictably?}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	%TODO: Order by Lastname?
	\author{Miriam Wagner\and
		Martin Breuer\and
		Moritz Werthebach\and
		Timo Bergerbusch\and
		Walter Schikowski}
	%
	\authorrunning{T. Bergerbusch et al.}
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
	%
	\maketitle              % typeset the header of the contribution
	%
	\begin{abstract}
		We analyze movements in the urban environment of the Columbian city Medellín. Each movement is given as spatiotemporal pattern of with additional information about the reason, means of transportation and the corresponding person like the socio-economic status (strata), the age and gender. 
		% Marías sentence
		The objective of the paper is twofold: try to find groups of movement patterns and see, whether they correspond to socio-economic status, and then, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
%		Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters. Due to bad results we introduce our preprocessing steps and switch to supervised learning.		
		Decision trees and neural networks did neither match our performance expectations, which leads to our conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.

>>>>

	%
	%	\title{Clustering Analysis of Mobility Data}
	\title{Are you moving predictably?}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	%TODO: Order by Lastname?
	\author{Miriam Wagner\and
		Martin Breuer\and
		Moritz Werthebach\and
		Timo Bergerbusch\and
		Walter Schikowski}
	%
	\authorrunning{T. Bergerbusch et al.}
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{RWTH Aachen, Templergraben 55, 52062 Aachen, Germany}
	%
	\maketitle              % typeset the header of the contribution
	%
	\begin{abstract}
		We analyze movements in the urban environment of the Columbian city Medellín. Each movement is given as spatiotemporal pattern with additional information about the reason of movment, means of transportation and the corresponding person concern the socio-economic status (strata), the age and gender. 
		% Marías sentence
		The objective of the paper is twofold: first we try to find groups of movement patterns and see, whether they correspond to socio-economic status, with this information, since we have the economic status of each person in the data set, we will apply supervised learning to classify patterns into socio-economic status.
%		Since in most cases we do not have information about the actual socio-economic status of persons we firstly try different unsupervised approaches to find natural clusters. Due to bad results we introduce our preprocessing steps and switch to supervised learning.		
		Neither decision trees nor neural networks match our performance expectations, which leads to the conclusion that we need more data and information about the data in order to find a proper social stratification and to predict the given socio-economic status accurately.

### 2018-07-15 dokumentation/theoreticalBackground.tex
koen hat korrigiert

	is a \textit{centroid based technique}, which means that each cluster is represented by a data point, that also is the centre of the cluster. A distance measure is then used to assign every remaining data object from the data set to the best fitting cluster. This is done according to its similarity to the centre of this cluster and its dissimilarity to the centres of any other cluster.

@@

	\begin{tabular}{r c l}
		\textbf{Retrieve} & : &  gives the data into the process \\
		\textbf{Generate ID} & : & creates an ID such that we can make the comparison\\
		 & & step at the end through joining the sets\\
		\textbf{Multiply} & : & creates two identical data sets\\
		\textbf{Select Attributes} & : &  troughs away the strata before the clustering step,\\
		& & everything except cluster and id after the clustering \\
		& & and just keeps id and strata for the join step\\
		\textbf{Clustering} & : &  runs the k-means clustering algorithm. The number of \\
		& & Clusters has to be fixed \\
		\textbf{Join} & : & For comparing the clustering result and the strata we \\
		& & join the two filtered data sets by the id \\
	\end{tabular}

@@

	Given two data points $x$ and $y$, each form a tuple of $v$ variables of arbitrary type, the similarity coefficient is given by
	\begin{equation}\label{eq1}
	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
	\end{equation} 
	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at the $k$-th entry in the data points $x$ and $y$. Note, that the definition of the score depends on the
	%Timo nervt
	type of the variable, as defined below. In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, if the two variables are comparable and to 0 if not. %PASS AUF
	For example, variables are not comparable, if values are undefined in the data points or the variable types do not match. Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$. Thus, the similarity coefficient in \hyperref[eq1]{(2)} can be interpreted as the average value of all similarity scores. 
	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation. As pointed above, 0 values are not considered as a match and even further, not considered to be comparable. Hence, the values result as in the table
		\begin{center}
			\begin{tabular}{l | c c c r}
				i & 1 & 1 & 0 & 0 \\
				j & 1 & 0 & 1 & 0 \\
				\hline
				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
			\end{tabular}
		\end{center}	
	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$, and 0, if they differ.

@@

\textit{PCA} uses an orthogonal transformation to convert a dataset with correlated datas into a dataset without correlations and decreased dimension. This is a helpful technique to have a better understanding of the data complexity as well as an improvment of time consumption in further steps.

@@

The Process does the following steps, like in \Cref{fig:PCA} visible:

\begin{tabular}{r c l}
	\textbf{Retrieve} & : & gives the data in the process.\\
	\textbf{Generate Id} & : & creates an idea for the join later\\
	\textbf{Select Attribute} & : & just gives the chosen attributes back\\
	\textbf{Nominal to Numerical} & : & changes the nominal data to numerical data, so \\
	&&that we can apply PCA in the next step \\
	\textbf{PCA} & : & applys the PCA reduction to the data\\
	&& threshold $0.95$ \\
	\textbf{Clustering} & : &  clusters the given data \\
	\textbf{Join} & : & joins the clustering outcome and the strata by the\\
	&& ID\\
\end{tabular}

@@

	\subsubsection{Decision Trees} are a good manner to figure out, which parts of the data set have the most influence on the decision. Therefore labelled data is needed and we have given the strata. We used again RapidMiner for building trees based on different data sets.

@@

	\begin{tabular}{r c l}
		\textbf{Retrieve} & : & includes the dataset\\
		\textbf{Select Attributes} & : & makes it possible to or have a look at grouped\\
		&& strata or normal strata\\
		\textbf{Set Role} & : & gives strata the label role, so that the decision tree\\
		&& has those as leafs\\
		\textbf{Multiply} & : & clones the data set \\
		\textbf{Decision Tree} &: & creates the decision tree\\
		\textbf{Apply Model} & : & creates the labelled data set for the \textbf{Performance} step\\
		\textbf{Performance} & : & gives the performance result of the created model\\
	\end{tabular}

>>>>

	is a \textit{centroid based technique}, which means that each cluster is represented by a data point, which at the same time is the centre of the cluster. A distance measure is then used to assign every remaining data object from the data set to the best fitting cluster. This is done according to its similarity to the centre of this cluster and its dissimilarity to the centres of any other cluster.

@@

	\begin{tabular}{r c l}
		\textbf{Retrieve} & : &  gives the data into the process \\
		\textbf{Generate ID} & : & creates an ID so we can make the comparison\\
		 & & step at the end through joining the sets\\
		\textbf{Multiply} & : & creates two identical data sets\\
		\textbf{Select Attributes} & : &  throws away the strata before the clustering step,\\
		& & everything except cluster and ID after the clustering \\
		& & and just keeps id and strata for the join step\\
		\textbf{Clustering} & : &  runs the k-means clustering algorithm. The number of \\
		& & Clusters has to be fixed \\
		\textbf{Join} & : & For comparing the clustering result and the strata we \\
		& & join the two filtered data sets by the ID \\
	\end{tabular}

@@

	Given two data points $x$ and $y$, each form a tuple of $v$ variables of arbitrary type, the similarity coefficient is given by
	\begin{equation}\label{eq1}
	S_{xy} = \sum_{k=1}^{v} s_{xy,k} / \sum_{k=1}^{v} \delta_{xy,k} 
	\end{equation} 
	where $s_{xy,k}$ denotes a \textit{score for the similarity} of the two variables at the $k$-th entry in the data points $x$ and $y$. Note, that the definition of the score depends on the
	%Timo nervt
	type of the variable, as defined below. In the divisor, $\delta_{xy,k}$ basically represents the possibility of comparing the two variables at index $k$, such that it evaluates to 1, if the two variables are comparable and to 0 if not. %PASS AUF
	For example, variables are not comparable, if values are undefined in the data points or the variable types do not match. Within this work, the dataset is complete, therefore, $\sum_{k=1}^{v} \delta_{xy,k}=v$. Thus, the similarity coefficient in \hyperref[eq1]{(2)} can be interpreted as the average value of all similarity scores. 
	With respect to the variable type, the similarity score $s_{xy,k}$ is defined as follows:\\
	\\ \textbf{Binary:} The score for binary variables is basically the result of an logical AND operation. As pointed out above, 0 values are not considered as a match such not even considered to be comparable. Hence, the values result as in the table
		\begin{center}
			\begin{tabular}{l | c c c r}
				i & 1 & 1 & 0 & 0 \\
				j & 1 & 0 & 1 & 0 \\
				\hline
				$s_{xy,k}$ & 1 & 0 & 0 & 0 \\
				$\delta_{xy,k}$ & 1 & 1 & 1 & 0
			\end{tabular}
		\end{center}	
	\textbf{Categorical:} The similarity score of categorical variables is 1, if the variables are completely identical in $x$ and $y$, and 0, if they differ.

@@

\textit{PCA} uses an orthogonal transformation to convert a dataset with correlated datas into a dataset without correlations and decreased dimension. This is a helpful technique to get a better understanding of the data complexity as well as decreasing the amount of time consumption in further steps.

@@

The Process does the following steps, visible in \Cref{fig:PCA}:

\begin{tabular}{r c l}
	\textbf{Retrieve} & : & imports the data in the process.\\
	\textbf{Generate Id} & : & creates an idea for the join later\\
	\textbf{Select Attribute} & : & gives the chosen attributes back\\
	\textbf{Nominal to Numerical} & : & changes the nominal data to numerical data, so \\
	&&that we can apply PCA in the next step \\
	\textbf{PCA} & : & applies the PCA reduction to the data\\
	&& threshold $0.95$ \\
	\textbf{Clustering} & : &  clusters the given data \\
	\textbf{Join} & : & joins the clustering outcome and the strata by the\\
	&& ID\\
\end{tabular}

@@

	\subsubsection{Decision Trees} are a good manner to figure out, which parts of the data set have the most influence on the decision. Therefore, labelled data is needed and we have given the strata. We again used RapidMiner for building trees based on different data sets.

@@

	\begin{tabular}{r c l}
		\textbf{Retrieve} & : & includes the dataset\\
		\textbf{Select Attributes} & : & makes it possible to have a look at grouped\\
		&& strata or normal strata\\
		\textbf{Set Role} & : & gives strata the label role, so that the decision tree\\
		&& has those as leafs\\
		\textbf{Multiply} & : & clones the data set \\
		\textbf{Decision Tree} &: & creates the decision tree\\
		\textbf{Apply Model} & : & creates the labelled data set for the \textbf{Performance} step\\
		\textbf{Performance} & : & gives the performance result of the created model\\
	\end{tabular}
